---
title: "Deep VARs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Deep VARs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(deepvars)
```

## Reduced-form VAR 

Recall the functional from of the reduced-form VAR($p$) with $K$ variables and $p$ lags (with $m \in [1,p]$) 

$$
\begin{aligned}
&& y_t&=A_1 y_{t-1} + A_2 y_{t-2} + ... + A_p y_{t-p} + u_t \\
\end{aligned}
$$

where $y_{t-m}$ are $(K \times 1)$ vectors , $A_m$ are $(K \times K)$ matrices of coefficients and $u_t$ is a $(K \times 1)$ vector of residuals. 

## Deep VAR

The deep VAR ...

## A simple benchmark

```{r example-data}
dt = data.table::data.table(deepvars::canada)
var_cols = colnames(dt)[2:ncol(dt)]
```

The underlying time series in levels look non-stationary 

```{r after-diff, fig.dim=c(7,4), fig.cap="Time series in differences."}
dt[,(var_cols) := lapply(.SD, function(i) c(0,diff(i))), .SDcols=var_cols]
chart_data = dt
chart_data = data.table::melt(chart_data, id.vars = "date")

ggplot2::ggplot(chart_data) +
  ggplot2::geom_line(ggplot2::aes(y=value, x=date)) +
  ggplot2::facet_wrap(variable ~ ., scales = "free_y") +
  ggplot2::theme_bw() +
  ggplot2::labs(
    x="Date",
    y="Value"
  )
```

Select lag length ...

```{r}
lags <- VAR_lag_select(dt[,.SD,.SDcols=var_cols])$p
```

Fit the two models ...

```{r}
# VAR
var_model <- vareg(dt, lags = lags)
# Deep VAR
deepvar_model <- deepvareg(dt, lags=lags, n_units = 100, epochs=150)
```

### Predictions

```{r, eval=FALSE}
y_true <- var_model$y_train
```

```{r, eval=FALSE, pred-var, fig.dim=c(7,4), fig.cap="Predictions from VAR."}
pred_var <- predict(var_model)
plot(pred_var, y_true = y_true)
```

```{r, eval=FALSE, pred-mlstm, fig.dim=c(7,4), fig.cap="Predictions from MLSTM."}
pred_deepvar <- predict(deepvar_model)
plot(pred_deepvar, y_true = y_true)
```

### MSE

```{r, eval=FALSE}
rmse_var <- rmse(var_model)[,model:="var"]
rmse_deep_var <- rmse(deepvar_model)[,model:="deepvar"]
tab_rmse <- rbind(rmse_var, rmse_deep_var)
tab_rmse <- data.table::dcast(tab_rmse, variable ~ model, value.var = "value")
knitr::kable(
  tab_rmse, 
  col.names = c("Variable", "Deep VAR", "VAR")
) 
```



