---
title: "Deep VARs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Deep VARs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)
```

```{r setup}
library(deepvars)
```

## Reduced-form VAR 

Recall the functional from of the reduced-form VAR($p$) with $K$ variables and $p$ lags (with $m \in [1,p]$) 

$$
\begin{aligned}
&& y_t&=A_1 y_{t-1} + A_2 y_{t-2} + ... + A_p y_{t-p} + u_t \\
\end{aligned}
$$

where $y_{t-m}$ are $(K \times 1)$ vectors , $A_m$ are $(K \times K)$ matrices of coefficients and $u_t$ is a $(K \times 1)$ vector of residuals. 

## Deep VAR

The deep VAR ...

## A simple benchmark

```{r example-data}
dt = data.table::data.table(deepvars::canada)
var_cols = colnames(dt)[2:ncol(dt)]
```

The underlying time series in levels look non-stationary 

```{r after-diff, fig.dim=c(7,4), fig.cap="Time series in differences."}
dt[,(var_cols) := lapply(.SD, function(i) c(0,diff(i))), .SDcols=var_cols]
chart_data = dt
chart_data = data.table::melt(chart_data, id.vars = "date")

ggplot2::ggplot(chart_data) +
  ggplot2::geom_line(ggplot2::aes(y=value, x=date)) +
  ggplot2::facet_wrap(variable ~ ., scales = "free_y") +
  ggplot2::theme_bw() +
  ggplot2::labs(
    x="Date",
    y="Value"
  )
```

Splitting data into a training and test sample ...

```{r}
train_test_split <- split_sample(dt)
train_data <- train_test_split$train_data
```

Select lag length ...

```{r}
lags <- VAR_lag_select(train_data[,.SD,.SDcols=var_cols])$p
```

Fit the two models ...

```{r}
# VAR
var_model <- vareg(train_data, lags = lags)
# Deep VAR
deepvar_model <- deepvareg(train_data, lags=lags, num_units = 100, epochs=100)
```

### Predictions

#### In-sample

```{r}
y_true <- var_model$y_train
```

```{r, pred-var, fig.dim=c(7,4), fig.cap="Predictions from VAR."}
pred_var <- predict(var_model)
plot(pred_var, y_true = y_true)
```

```{r, pred-mlstm, fig.dim=c(7,4), fig.cap="Predictions from MLSTM."}
pred_deepvar <- predict(deepvar_model)
plot(pred_deepvar, y_true = y_true)
```

#### Out-of-sample

```{r, warning=FALSE}
X_test <- prepare_test_data(train_test_split, lags=lags)$X_test
y_test <- prepare_test_data(train_test_split, lags=lags)$y_test
```


```{r, out-pred-var, fig.dim=c(7,4), fig.cap="Out-of-sample predictions from VAR."}
pred_var <- predict(var_model, X=X_test)
plot(pred_var, y_true = y_test)
```

```{r, out-pred-dvar, fig.dim=c(7,4), fig.cap="Out-of-sample predictions from DVAR."}
pred_dvar <- predict(deepvar_model, X=X_test)
plot(pred_dvar, y_true = y_test)
```

### Cumulative loss

```{r, cum-loss, fig.dim=c(7,4), fig.cap="Comparison of cumulative loss for the two models."}
cum_loss_var <- cum_loss(var_model)$cum_loss[,type:="var"]
cum_loss_dvar <- cum_loss(deepvar_model)$cum_loss[,type:="deepvar"]
dt_plot <- rbind(cum_loss_dvar, cum_loss_var)
ggplot2::ggplot(data=dt_plot, ggplot2::aes(x=date, y=value, colour=type)) +
  ggplot2::geom_line() +
  ggplot2::facet_wrap(~variable, scales = "free_y") +
  ggplot2::scale_color_discrete(name="Model:") +
  ggplot2::labs(
      x="Date",
      y="Squared error"
    )
```


### RMSE

```{r}
rmse_var <- rbind(
  rmse(var_model)[,model:="var"][,sample:="train"],
  rmse(var_model, X=X_test, y=y_test)[,model:="var"][,sample:="test"]
)
rmse_dvar <- rbind(
  rmse(deepvar_model)[,model:="deepvar"][,sample:="train"],
  rmse(deepvar_model, X=X_test, y=y_test)[,model:="deepvar"][,sample:="test"]
)
tab_rmse <- rbind(rmse_var, rmse_dvar)
tab_rmse <- data.table::dcast(tab_rmse, sample + variable ~ model, value.var = "value")
knitr::kable(
  tab_rmse, 
  col.names = c("Sample", "Variable", "DVAR", "VAR"),
  digits = 2
) 
```

### Forecasts

```{r, fcst-var, fig.dim=c(7,4), fig.cap="Forecasts from VAR."}
fcst_var <- forecast(var_model, n.ahead = 50)
plot(fcst_var)
```

```{r, fcst-dvar, fig.dim=c(7,4), fig.cap="Forecasts from Deep VAR."}
fcst_dvar <- forecast(deepvar_model, n.ahead = 50)
plot(fcst_dvar)
```

