---
title: "Structural VARs (`deepvars`)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Structural VARs (`deepvars`)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(deepvars)

# Dependencies
library(tidyr)
library(data.table)
library(ggplot2)
library(expm)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo=F,
  warning = F, 
  message = F
)
```

## Structural VAR

The structural VAR model accounts for contemporaneous realtionships between variables entering the model. Mathematically, it is the reduced-form model pre-multiplied by $B_0$, a matrix capturing the contemporaneous relationships:

$$
\begin{aligned}
&& y_t&=A_1 y_{t-1} + A_2 y_{t-2} + ... + A_p y_{t-p} + u_t && \text{(Reduced)} \\
&& B_0y_t&=B_0A_1 y_{t-1} + B_0A_2 y_{t-2} + ... + B_0A_p y_{t-p} + \varepsilon_t \\
&& B_0y_t&=B_1 y_{t-1} + B_2 y_{t-2} + ... + B_p y_{t-p} + \varepsilon_t && \text{(Structural)} \\
\end{aligned}
$$

By convention the covariance matrix of the structural errors $\varepsilon_t$ is normalised to $\Sigma_{\varepsilon} = I_k$ such that $\Sigma_u=B_0^{-1}{B_0^{-1}}'$. Note that $\frac{(k+1)k}{2}$ elements of $\Sigma_u$ can be uniquely estimated in the reduced form VAR (Gottschalk 2001), since off-diagonal elements in the lower and upper triangle are duplicates of each other. Thus, one needs to impose restrictions on the $\frac{(k-1)k}{2}$
remaining parameters of $B_0$ for the model to be just identified.

We will consider the following example data set:

```{r example-data}
dt = data.table::data.table(deepvars::canada)
var_cols = colnames(dt)[2:ncol(dt)]
print(head(dt))
dt[,(var_cols) := lapply(.SD, function(i) c(0,diff(i))), .SDcols=var_cols]
lag_selection = VAR_lag_select(dt[,.SD,.SDcols=var_cols])
lags = lag_selection$p
varresult = vareg(dt, lags = lags)
```

### Impulse response functions

#### Bootstrapped confidence-intervals

```{r, fig.dim=c(7,4)}
irf_output = irf(varresult, imp = 1)
```

### Variance decompositions

#### Forecast error variance decomposition (FEVD)

```{r, fig.dim=c(7,4)}
fevd_output = fevd(varresult)
fevd_output$plot
```


#### Historical decompositon

```{r, eval=F, fig.dim=c(7,4)}
# Historical decomposition as in Kilian and Luetkepohl ----
csvs = list.files("data/kilian_lee")
kilian_lee = lapply(1:length(csvs), function(i) {
  fread(sprintf("data/kilian_lee/%s", csvs[i]))
})
names(kilian_lee) = gsub(".csv","",csvs)
A_comp = as.matrix(kilian_lee$A)
B_0 = solve(kilian_lee$B0inv)
lag = 24
dt = kilian_lee$y
N = nrow(dt)-lag
K = ncol(dt)
u = as.matrix(kilian_lee$Uhat)

var_names = c("a", "b", "Oil price", "d")
colnames(dt) = var_names

# Step 1 - compute structural coefficient matrices: ----
Phi = red_ir(lag, K, A_comp, const, n.ahead=N)
Theta = compute_Theta(Phi, B_0)

# Step 2 - compute the structural shocks: ----
w = u %*% t(B_0)

# Step 3 - match structural shocks with appropriate impulse response weight: ----
y_hat_dt = data.table::rbindlist(
  lapply(1:N, function(t) { # loop over time periods
    terms_to_sum = lapply(1:t, function(s) {
      sapply(1:K, function(k) Theta[[s]][k,] * w[t-s+1,])
    })
    y_hat_t = Reduce(`+`, terms_to_sum)
    colnames(y_hat_t) = var_names
    y_hat_t = data.table::data.table(t(y_hat_t))[,c("k","t"):=.(var_names, t)]
    return(y_hat_t)
  })
)[order(k,t)]

# Step 4 - tidy up: ----
y_hat_dt[,actual:=detrend(dt[,k]),by=k]
y_hat_dt[,fitted:=rowSums(.SD),.SDcols=var_names]
y_hat_dt[,residual:=actual-fitted]
```

```{r, fig.dim=c(7,4)}
hd(varresult = varresult)
```

